Ionization of molecules by multicharged bared ions by using the stoichiometric model

### 2019-09-18 NewMonti25.tex
added experiments on DNA mols
changed colors

\begin{figure}[H]
\centering
%\includegraphics[width=\textwidth]{figuras/atomicscaling.eps}
\includegraphics[width=0.7\textwidth]{figuras/Fig_finales/Fig1.eps}
\caption{Scaled total ionization cross section of six atomic targets 
by impact of six projectiles.}
\label{eq:atomscaling}
\end{figure} 
In Figure 1, we report our total ionization results for the six 
essential elements by the impact of six different projectiles: 
antiprotons, H$^{+}$, He$^{+2}$, Be$^{+4}$, C$^{+6}$, and O$^{+8}$. 
To reduce the resulting 36 magnitudes into a single consistent 
figure, we considered the fact that in the first Born approximation
the ionization cross-section scales with the square of the projectile 
charge, $Z^{2}$. The values of the impact energies considered 
range between 0.1 to 10 MeV/amu, where the CDW is supposed 
to hold. In fact, for the highest projectile charges the minimum 
impact energy where the CDW is expected to be valid could be 
higher than 100 KeV. Moreover, we corroborated that the first Born 
approximation provides quite reliable results only for energies higher 
than a couple of MeV/amu. We use the same color to indicate the 
projectile charge throughout the figures of this work: dashed-red, 
solid-red, blue, magenta, olive and orange for antiprotons, 
H$^{+}$, He$^{+2}$, Be$^{+4}$, C$^{+6}$, and O$^{+8}$, respectively. 
Notably, there is no complete tabulation of ionization of atoms by 
the impact of multicharged ions. We hope that the ones presented 
in this article will be of help for future works.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Ionization of Molecules}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{The stoichiometric model}

Lets us consider a molecule $M$ composed by $n_{\alpha}$ atoms of the
element $\alpha$, the SSM describes the total ionization cross section 
of the molecule $\sigma_{M}$ as a simple sum of ionization cross 
sections of the isolated atoms $\sigma_{\alpha}$, 
\begin{equation}
 \sigma_{M}=\sum\limits_{\alpha}n_{\alpha}\sigma_{\alpha}\,.  
 \label{eq:sumion}
\end{equation}
We divided sixteen molecular targets of our interest in three families. 
The classification defined is given in Table~\ref{tab:families}.
\begin{table}
\begin{center}
\begin{tabular}{|p{0.06\textwidth}|p{0.56\textwidth}|}
\hline
 CH  & CH$_4$, C$_2$H$_2$, C$_2$H$_4$, C$_2$H$_6$, C$_6$H$_6$ \\
\hline
 CHN & C$_5$H$_5$N, C$_4$H$_4$N$_2$, C$_2$H$_7$N, CH$_5$N \\
\hline
 \multirow{5}{*}{DNA} 
     & C$_5$H$_5$N$_5$ (adenine), C$_4$H$_5$N$_3$O (cytosine), \\
     & C$_5$H$_5$N$_5$O (guanine), C$_5$H$_6$N$_2$O$_2$ (thymine), \\
     & C$_4$H$_4$N$_2$O$_2$ (uracil), C$_4$H$_8$O (THF), \\
     & C$_5$H$_{10}$O$_5$P (DNA backbone), \\
     & C$_{20}$H$_{27}$N$_7$O$_{13}$P$_2$ (dry DNA) \\
\hline
\end{tabular}
\caption{Seventeen molecular targets of our interest classified in three 
families.}
\label{tab:families}
\end{center}
\end{table}

In Figure 2, we report the total ionization cross sections by the 
impact of multicharged ions for adenine, cytosine, guanine and thymine. 
For adenine, the agreement with the experimental data 
available (circles)~\cite{iriki2011} is quite good in our range of validity. 
To the best of our knowledge, there are not experimental data on 
ion--collision ionization for the rest of the molecules. We included 
electron impact measurements (right--oriented triangles)~\cite{mozejko2003} 
for incident energies greater than 500keV, were good agreement is found 
with proton ionization for all the molecules. 
%In fact, recent models for electron impact in complex biomolecules 
%have been under progress~\ref{tan2018}.

\begin{figure}[H]
\centering
\includegraphics[width=0.7\textwidth]{figuras/Fig_finales/Fig2.eps}
\caption{Present ionization cross section by impact of six projectiles, 
and experimental values for proton and electron ionization.}
\label{eq:crossDNA}
\end{figure}

Our reports for uracil, DNA backbone, pyrimidine and THF are given in 
Figure 3. For uracil, we have good agreement with the experimental 
proton impact measurements by Itoh~{\it et al.}~\cite{itoh2013}. 
However, for the same target, our theory fails by a factor of two for 
the experimental ionization values by the impact of C$^{+4}$~(square symbols), 
O$^{+6}$~(circle symbols) and O$^{+8}$~(triangle--down symbols) 
ions~\cite{agnihotri2012,agnihotri2013}. 
Nonetheless, it should be stated that our theoretical results coincide 
with calculations by Champion, Rivarola and 
collaborators~\cite{champion2012,agnihotri2012}, which may indicate a 
possible misstep of the experiments. 
For pyrimidine, we show comparison of our results with experimental 
proton~\cite{wolff2014} and electron~\cite{bug2017} ionization. 
The electron impact measurements agree with our calculations for energies
higher than 500keV. Unexpectedly, the proton impact cross sections are
significantly lower than our findings. 
The THF molecule results are compared with proton~\cite{wang2016} and 
electron~\cite{wolf2019,bug2017,fuss2009} impact measurements, showing
overall good agreement with our theory.

\begin{figure}[H]
\centering
\includegraphics[width=0.7\textwidth]{figuras/Fig_finales/Fig3.eps}
\caption{Present ionization cross section by impact of six projectiles, 
and experimental values for proton and electron ionization.}
\label{eq:crossDNA}
\end{figure}

\subsection{Scaling rule}

The first attempt to develop a comprehensive but straightforward 
phenomenological model for electron ejection from large molecules was 
proposed by Toburen {\it et al.}~\cite{toburen1975,toburen1976}. 
The authors found it convenient to scale the experimental ionization 
cross section in terms of the number of outer or weakly bound valence 
electrons (total number of electrons minus the inner-shell ones). 
We introduce the cross section per weakly bound electron or ionization 
cross section per valence electron, $\sigma_{e}$, as
\begin{equation}
\sigma_{e}=\frac{\sigma_{M}}{N_{M}}=\frac{\sum\limits_{\alpha}
n_{\alpha}\sigma_{\alpha}}{\sum\limits_{\alpha}n_{\alpha}\nu_{\alpha}}
=\sigma_{e}(v)\,, 
\label{27} 
\end{equation}
where $\nu_{\alpha}$ is the number of weakly bound electrons of the atom
$\alpha$, given by
\begin{equation}
\nu_{\alpha}=\left\{ 
\begin{array}{ll}
4, & \text{for C,} \\ 
5, & \text{for N and P,} \\ 
6, & \text{for O and S,}
\end{array}\right.
\label{eq:nelec} 
\end{equation}
and $\nu_{\text{H}}=1$. The Toburen rule can be stated by saying that 
$\sigma_{e}$ is a \textit{universal} parameter independent on the 
molecule, which depends solely on the impact velocity. 
In Ref.~\cite{toburen1976}, it was found experimentally that for proton 
impact on some simple molecules, $\sigma_{e}$ results
\begin{equation}
\begin{array}{cccc}
E\text{(MeV)}               & 0.25 & 1.00 & 2.00 \\ 
\sigma_{e} (10^{-16}cm^{2}) & 0.39 & 0.17 & 0.11
\end{array}
\label{30}
\end{equation}
with an estimated error of about 20\%. A similar ratio was found in 
Ref.~\cite{itoh2013} for proton impact on uracil and adenine.

\begin{figure}[t!]
\includegraphics[width=\textwidth]{figuras/molscaling85.eps}
\caption{Ionization cross section per weakly bound electron using
(a) Eq.~(\ref{eq:nelec}) and (b) Eq.~(\ref{eq:scalingCDW}).}
\label{fig:newscaling}
\end{figure}
Figure 4a shows $\sigma_{e}$ calculated with the CDW as a function of 
the impact energy for different projectile charges computed with the 
SSM for the seventeen molecular targets displayed in 
Table~\ref{tab:families}. The universality with $Z$ is the one provided 
by Born approximation, i.e., $\sigma _{e}(Z)=Z^{2}\sigma_{e}(Z=1)$, 
and it holds for large impact velocities, as shown Figure 4a.
Of course, for lower impact velocities, the CDW breaks the behavior of 
the $Z^{2}$ rule. Although the Toburen rule holds for high energies, 
its performance is still not satisfactory: the universal band is quite 
broad. The departure of our theoretical 
results from the Toburen rule can be easily explained by the fact that 
our atomic cross sections $\sigma_{\alpha}$ behave differently for 
each atom in terms of the projectile charges. By inspecting Figure 1, 
one can easily see that the rule $\sigma_{\alpha}/\nu_{\alpha}\sim
\sigma_{e}$ is not well satisfied by the CDW. A much better general 
atomic rule is given by
\begin{equation}
\sigma_{e}^{\text{CDW}}=\frac{\sum\limits_{\alpha}
n_{\alpha}\sigma_{\alpha}^{\text{CDW}}}{\sum\limits_{\alpha}n_{\alpha}
\nu_{\alpha}^{\text{CDW}}},
\label{32} 
\end{equation}
where 
\begin{equation}
\nu_{\alpha }^{\text{CDW}} \sim\left\{ 
\begin{array}{ll}
4,\ \ \  & \text{for C, N, and O,} \\ 
4.5,\ \  & \text{for P and S}
\end{array}
\right. 
\label{eq:scalingCDW}
\end{equation}
and, obviously, $\nu_{\text{H}}^{\text{CDW}}=1$. The cross sections 
$\sigma_{e}^{\text{CDW}}$ are plotted in Figure~4b. A much better sharp 
band is observed, especially at high energy, where the theory is 
expected to work. It will be interesting for experiments to cross-check 
this theoretical prediction.

\subsection{Emitted electron energies}

@@

Figure 5 shows $\overline{E}_{\alpha}$ for six elements from 
Table~\ref{tab:families}. The range of impact velocities was shorten up 
to $v=10$ a.u. due to numerical limitations in the spherical harmonics 
expansion. In our theoretical treatment, we expand our final continuum 
wave function as per usual,
\begin{equation}
\psi_{\overrightarrow{k}}^{-}(\overrightarrow{r})=\sum_{l=0}^{l_{\max
}}\sum_{m=-l}^{l}R_{kl}^{-}(r)Y_{l}^{m}(\widehat{r})Y_{l}^{m^{\ast }}
(\widehat{k})\,.
\label{60}
\end{equation}
We are confident with our calculations up to $l_{\max}\sim 30$. 
As the impact velocity $v$ increases, so do $\langle E_{\alpha}\rangle$
and $l_{\max}$, which results in the inclusion of very oscillatory 
functions in the integrand. Furthermore, the integrand of
$\langle E_{\alpha}\rangle$ includes the kinetic energy $E$
(see Eq.~(\ref{40})), which cancels the small energy region and 
reinforces the large values, making the result more sensible to large
angular momenta. Regardless, for $v>10$ a.u., the first Born 
approximation holds.

In Figure 5, we estimate $\overline{E}_{\alpha}$ in the 0.5--2 a.u.
velocity range, or equivalently from 15 to 50 eV, for all the targets.
Our results agree with the experimental findings~\cite{surdutovic2018}. 
The dependence of the mean energy value with the projectile charge $Z$ 
is surprisingly sensible, which can duplicate the proton results. 
This effect can be attributed to the depletion caused by the 
multicharged ions to the yields of low energy electrons. In the high 
electron energy regime, the CDW falls on the simple first Born 
approximation, surviving the $Z^{2}$ law. Then, the ratio in 
Eq.~(\ref{40}) cancels out and $\overline{E}_{\alpha}$ becomes a 
universal value independent on $Z$.

@@

\subsection{Emitted electron angles}

As mentioned before, secondary electrons contribute to the total damage. 
Then, not only is the ejection energy important but also the angle 
of emission. Once again, we can consider the single differential cross 
section in terms of the ejected electron solid angle $\Omega$, 
$d\sigma_{\alpha,nl}/d\Omega$, to be expressed as a distribution function, 
and the mean angle $\overline{\theta}_{\alpha}$ can be defined as
\begin{equation}
\overline{\theta}_{\alpha}=\frac{\langle\theta_{\alpha}\rangle}
{\langle 1\rangle}=\frac{1}{\sigma_{\alpha}}\sum\limits_{nl}
\int d\Omega\,\theta\,\frac{d\sigma_{\alpha,nl}}{d\Omega}
\end{equation}
Figure 6 shows $\overline{\theta}_{\alpha}$ for our six elements of 
interest. A new important dependence of $\overline{\theta}_{\alpha}$ 
with $Z$ is observed. It is a general belief~\cite{Rudd1992} that the 
angular dispersion of emitted electrons are nearly isotropic. This effect 
is caused by the insignificant angular anisotropy of sub-50-eV yield. 
A typical value for the ejection angle considered in the literature is 
$\overline{\theta}_{\alpha}\sim 70\degree$~\cite{surdutovic2018}, and 
it is quite correct in the range of validity of the first Born 
approximation for any target. But, when a distorted wave approximation 
is used, $\overline{\theta}_{\alpha}$ decreases substantially with $Z$ 
in the intermediate energy region, as observed in Figure 6. 
For example, for C$^{+6}$ impact, the Bragg peak occurs at 0.3 MeV/amu, 
where $\overline{\theta}_{\alpha}$, computed with the CDW method, is 
about half of the value obtained with the first Born approximation. 
This correction should close the damage to the forward direction.

@@

\subsection{A modified stoichiometric model}

@@

\bibitem{mozejko2003}
P. Mozejko and L. Sanche, Radiat Environ. 
Biophys \textbf{42}, 201 (2003).

>>>>

In Fig.~\ref{fig:atomscaling}, we report our total ionization results 
for the six essential elements by the impact of six different projectiles: 
antiprotons, H$^{+}$, He$^{+2}$, Be$^{+4}$, C$^{+6}$, and O$^{+8}$. 
To reduce the resulting 36 magnitudes into a single consistent 
figure, we considered the fact that in the first Born approximation
the ionization cross-section scales with the square of the projectile 
charge, $Z^{2}$. The values of the impact energies considered 
range between 0.1 to 10 MeV/amu, where the CDW is supposed 
to hold. In fact, for the highest projectile charges the minimum 
impact energy where the CDW is expected to be valid could be 
higher than 100 KeV. Moreover, we corroborated that the first Born 
approximation provides quite reliable results only for energies higher 
than a couple of MeV/amu. We use the same color to indicate the 
projectile charge throughout the figures of this work: dashed-red, 
solid-red, blue, magenta, olive and orange for antiprotons, 
H$^{+}$, He$^{+2}$, Be$^{+4}$, C$^{+6}$, and O$^{+8}$, respectively. 
The charge state of the projectiles are also given in the figures.
Notably, there is no complete tabulation of ionization of atoms by 
the impact of multicharged ions. We hope that the ones presented 
in this article will be of help for future works.

\begin{figure}[t!]
\centering
\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/Fig1.eps}
\caption{Scaled total ionization cross section of six atomic targets 
by impact of multicharged projectiles.}
\label{fig:atomscaling}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Ionization of Molecules}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The stoichiometric model}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Lets us consider a molecule $M$ composed by $n_{\alpha}$ atoms of the
element $\alpha$, the SSM describes the total ionization cross section 
of the molecule $\sigma_{M}$ as a simple sum of ionization cross 
sections of the isolated atoms $\sigma_{\alpha}$, 
\begin{equation}
 \sigma_{M}=\sum\limits_{\alpha}n_{\alpha}\sigma_{\alpha}\,.  
 \label{eq:sumion}
\end{equation}
We divided sixteen molecular targets of our interest in three families. 
The classification defined is given in Table~\ref{tab:families}.

\begin{table}[H]
\begin{center}
\begin{tabular}{|p{0.06\textwidth}|p{0.75\textwidth}|}
\hline
 CH  & CH$_4$ (methane), C$_2$H$_2$ (acetylene), \\
     & C$_2$H$_4$ (ethene), C$_2$H$_6$ (ethane), C$_6$H$_6$ (benzene) \\
\hline
 CHN & C$_5$H$_5$N (pyridine), C$_4$H$_4$N$_2$ (pyrimidine), \\
& C$_2$H$_7$N 
(dimenthylamine), CH$_5$N (monomethylamine) \\
\hline
 \multirow{4}{*}{DNA} 
     & C$_5$H$_5$N$_5$ (adenine), C$_4$H$_5$N$_3$O (cytosine), \\
     & C$_5$H$_5$N$_5$O (guanine), C$_5$H$_6$N$_2$O$_2$ (thymine), \\
     & C$_4$H$_4$N$_2$O$_2$ (uracil), C$_4$H$_8$O (THF), \\
     & C$_5$H$_{10}$O$_5$P (DNA backbone), C$_{20}$H$_{27}$N$_7$O$_{13}$P$_2$ (dry DNA) \\
\hline
\end{tabular}
\caption{Molecular targets of our interest classified in three families.}
\label{tab:families}
\end{center}
\end{table}

We report the total ionization cross sections by the impact of 
multicharged ions for adenine, cytosine, guanine and thymine in 
Fig.~\ref{fig:crossDNA_1}. For adenine, the agreement with the 
experimental data available (circles)~\cite{iriki2011} is very good 
in our range of validity. To the best of our knowledge, there are not 
experimental data on ion--collision ionization for the rest of the 
molecules. We included electron impact measurements (right--oriented 
triangles)~\cite{rahman2016} with the corresponding equivelocity 
conversion for incident energies higher than 500~keV. In this region, 
the proton and electron cross section should agree. Although the electron 
impact measurements are above our findings for all the molecular targets, 
it is worth stating that our results agree very well with other 
theoretical predictions~\cite{mozejko2003,tan2018}. The symbols used to 
make reference to the proton and electron experimental data will be kept 
for subsequent figures.

\begin{figure}[t!]
\centering
\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig2.eps}
\caption{Present ionization cross section by impact of six projectiles, 
and experimental values for proton and electron ionization.}
\label{fig:crossDNA_1}
\end{figure}

Our calculations for uracil, DNA backbone, pyrimidine and THF are given in 
Fig.~\ref{fig:crossDNA_2}. 
For uracil, we have good agreement with the experimental 
proton impact measurements (triangle--up) by Itoh~{\it et al.}~\cite{itoh2013}. 
However, for the same target, our theory fails by a factor of two for 
the experimental ionization values by the impact of 
C$^{+4}$~(squares), O$^{+6}$~(circles) and O$^{+8}$~(triangles--down) 
ions~\cite{agnihotri2012,agnihotri2013}.
Nonetheless, it should be stated that our theoretical results coincide 
with calculations by Champion, Rivarola and 
collaborators~\cite{champion2012,agnihotri2012}, which may indicate a 
possible misstep of the experiments. 
For pyrimidine, we show comparison of our results with experimental data
for proton (circles with dash)~\cite{wolff2014} and electron
(stars)~\cite{bug2017} ionization. The electron impact measurements 
agree with our calculations for energies higher than 500keV. 
Unexpectedly, the proton impact cross sections are significantly lower 
than our findings. 
The THF molecule results are compared with proton (squares)~\cite{wang2016} 
and electron~\cite{bug2017,wolf2019,fuss2009} impact measurements, showing
overall good agreement with our theory.

\begin{figure}[t!]
\centering
\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/Fig3.eps}
\caption{Present ionization cross section by impact of six projectiles, 
and experimental values for proton and electron ionization.}
\label{fig:crossDNA_2}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Scaling rule}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Toburen numbers}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The first attempt to develop a comprehensive but straightforward 
phenomenological model for electron ejection from large molecules was 
proposed by Toburen {\it et al.}~\cite{toburen1975,toburen1976}. 
The authors found it convenient to scale the experimental ionization 
cross section in terms of the number of active or weakly--bound valence 
electrons $n_e$ (i.e. total number of electrons minus the K--shell).
Following Toburen, we can define the ionization cross section per weakly 
bound electron, $\sigma_{e}^T$, as
\begin{equation}
\sigma_{e}^T=\frac{\sigma_{M}}{n_e}=\frac{\sum\limits_{\alpha}
n_{\alpha}\sigma_{\alpha}}{\sum\limits_{\alpha}n_{\alpha}\nu_{\alpha}^T}
=\sigma_{e}^T(v)\,, 
\label{27} 
\end{equation}
where $\nu_{\alpha}^T$ are the Toburen numbers for each $\alpha$ atom 
is given by
\begin{equation}
\nu_{\alpha}^T=\left\{ 
\begin{array}{ll}
1, & \text{for H,} \\
4, & \text{for C,} \\ 
5, & \text{for N and P,} \\ 
6, & \text{for O and S}\,.
\end{array}\right.
\label{eq:nelec} 
\end{equation} 
The Toburen rule can be stated by saying that 
$\sigma_{e}$ is a \textit{universal} parameter independent on the 
molecule, which depends solely on the impact velocity, and holds for 
high impact energies (i.e. 0.25--5 Mev/amu).
%In Ref.~\cite{toburen1976}, it was found experimentally that for proton 
%impact on some simple molecules, $\sigma_{e}$ results
%\begin{equation}
%\begin{array}{cccc}
%E\text{(MeV)}               & 0.25 & 1.00 & 2.00 \\ 
%\sigma_{e} (10^{-16}cm^{2}) & 0.39 & 0.17 & 0.11
%\end{array}
%\label{30}
%\end{equation}
%with an estimated error of about 20\%. A similar ratio was found in 
%Ref.~\cite{itoh2013} for proton impact on uracil and adenine. 
These $\nu_{\alpha}^T$ can be interpreted as the number of active 
electrons in the collision. Of course, at very high energies also de 
K--shell electrons will be ionized and these numbers will be different.
A similar dependence with the number of weakly bound electrons was 
found in Ref.~\cite{itoh2013} for proton impact on uracil and adenine.

\begin{figure}[t!]
\centering
\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/Fig4.eps}
\caption{Scaled ionization cross section per weakly bound electron using
(a)~the Toburen numbers $\nu_{\alpha}^T$, and (b) our proposed numbers
$\nu_{\alpha}^{\text{CDW}}$.}
\label{fig:newscaling}
\end{figure}

Our CDW results for $\sigma_{e}^T$ are shown in 
Fig.~\ref{fig:newscaling}a as a function of 
the impact energy for different projectile charges computed with the 
SSM for the molecular targets from Table~\ref{tab:families}. 
The universality with $Z$ is the one provided 
by Born approximation, i.e., $\sigma_{e}^T(Z)=Z^{2}\sigma_{e}^T(Z=1)$, 
and it holds for large impact velocities, as shown 
Fig.~\ref{fig:newscaling}a.
Of course, for lower impact velocities, the CDW breaks the behavior of 
the $Z^{2}$ rule. Although the Toburen scaling holds for high energies, 
its performance is still not satisfactory: the universal band is quite 
broad.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{New scaling}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The departure of our theoretical 
results from the Toburen rule can be easily understood %explained by the fact that 
%our atomic cross sections $\sigma_{\alpha}$ behave differently for 
%each atom in terms of the projectile charges. By 
by inspecting Fig.~\ref{fig:atomscaling}. It can be noted that the 
rule $\sigma_{\alpha}/\nu_{\alpha}^T\sim \sigma_{e}^T$, approximatelly 
constant, is not well satisfied by the CDW. 
For example, Fig.~\ref{fig:atomscaling} shows that the cross sections
for O are actually very similar to the cross sections for C, suggesting 
4 active electrons in O instead of 6. In the same way, the number of
active electrons for N, P and S are also different from the 
$\nu_{\alpha}^T$ of Eq.~(\ref{eq:nelec}).

Based on the CDW results, we propose a new scaling,
\begin{equation}
\sigma_{e}=\frac{\sigma_M}{\sum\limits_{\alpha}n_{\alpha}
\nu_{\alpha}^{\text{CDW}}},
\label{32} 
\end{equation}
where $\sigma_M$ follows Eq.~(\ref{eq:sumion}), and 
$\nu_{\alpha}^{\text{CDW}}$ are the new active--electron numbers 
per atom obtained from the CDW ionization cross sections for 
different ions in H, C, N, O, P , and S targets. 
Our proposed active--electron numbers are
\begin{equation}
\nu_{\alpha }^{\text{CDW}} \sim\left\{ 
\begin{array}{ll}
1, & \text{for H,} \\
4, & \text{for C, N, and O,} \\ 
4.5, & \text{for P and S}\,.
\end{array}
\right. 
\label{eq:scalingCDW}
\end{equation}

The new scaled cross sections $\sigma_{e}$ are plotted in Figure~4b. 
A much better sharp band is observed, especially for impact energies 
$E=(0.5-8)$~MeV/amu for $Z=1$ and $Z=2$, and $E=(2.5-8)$~MeV/amu for 
$Z>2$. In fact, the experimental data for ionization of adenine, 
cytosine, thymine, guanine, uracil, pyrimidine and THF by proton impact 
seems to corroborate the new scaling. We also included the electron impact
ionization measurements from Figs.~\ref{fig:crossDNA_1} and 
\ref{fig:crossDNA_2}. It will be interesting to cross--check for 
experiments with higher projectile charge states.

\begin{table}[H]
\begin{center}
\begin{tabular}{|ll|ll|ll|}
\hline
 Molecule & $n_e$ &Molecule         & $n_e$ & Molecule             & $n_e$ \\
\hline
 H$_2$    & 2  & C$_2$H$_7$N         & 19    & C$_4$H$_5$N$_3$O     & 37    \\
 H$_2$O   & 6  & C$_4$H$_8$O         & 28    & C$_5$H$_6$N$_2$O$_2$ & 42    \\
 NH$_3$   & 7  & C$_4$H$_4$N$_2$     & 28    & C$_5$H$_5$N$_5$      & 45    \\
 CH$_4$   & 8  & C$_6$H$_6$          & 30    & C$_5$H$_5$N$_5$O     & 49    \\
 CH$_5$N  & 13 & C$_4$H$_4$N$_2$O$_2$& 36    & C$_5$H$_{10}$O$_5$P  & 54.5  \\
 \hline
\end{tabular}
\caption{New scaling numbers for some molecular targets of biological interest.}
\label{nn}
\end{center}
\end{table}

By using Eq.~(\ref{eq:scalingCDW}), the number of active electrons 
$n_e=\sum_{\alpha} n_{\alpha} \nu_{\alpha}^{\text{CDW}}$ can be redefined. 
We give new values in Table~\ref{nn} for some molecules of interest.
These values are very different from the ones proposed by Toburen and used by 
other authors~\cite{itoh2013}.
Moreover, an alternative way of showing the scaling can be attained by 
plotting the ionization cross sections of molecules as a function of the 
number of active electrons from Table~\ref{nn}. Our findings are displayed 
in Fig.~\ref{fig:recta} for impact energies 0.5, 1 and 2 MeV. As can be 
noted, the computed CDW ionization cross sections for all the molecules  
show a linear dependence with the number of electrons from Table~\ref{nn}. 
We obtain similar results even for $E=10$~MeV. The comparison with the 
experimental data available shows very nice agreement, from the smallest 
molecules, H$_2$, H$_2$O and CH$_4$, up to the most complex ones, like 
adenine or guanine. In some cases, the experimental data were interpolated.
Care was taken in such procedure.

\begin{figure}[t!]
\centering
\includegraphics[width=0.75\textwidth]{figuras/fig_recta.eps}
\caption{.}
\label{fig:recta}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Emitted electron energies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{figure}[t!]
\centering
\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/Fig5.eps}
\caption{.}
\label{fig:emittedener}
\end{figure}

Figure~\ref{fig:emittedener} shows $\overline{E}_{\alpha}$ for six elements from 
Table~\ref{tab:families}. The range of impact velocities was shorten up 
to $v=10$ a.u. due to numerical limitations in the spherical harmonics 
expansion. In our theoretical treatment, we expand our final continuum 
wave function as per usual,
\begin{equation}
\psi_{\overrightarrow{k}}^{-}(\overrightarrow{r})=\sum_{l=0}^{l_{\max
}}\sum_{m=-l}^{l}R_{kl}^{-}(r)Y_{l}^{m}(\widehat{r})Y_{l}^{m^{\ast }}
(\widehat{k})\,.
\label{60}
\end{equation}
We are confident with our calculations up to $l_{\max}\sim 30$. 
As the impact velocity $v$ increases, so do $\langle E_{\alpha}\rangle$
and $l_{\max}$, which results in the inclusion of very oscillatory 
functions in the integrand. Furthermore, the integrand of
$\langle E_{\alpha}\rangle$ includes the kinetic energy $E$
(see Eq.~(\ref{40})), which cancels the small energy region and 
reinforces the large values, making the result more sensible to large
angular momenta. Regardless, for $v>10$ a.u., the first Born 
approximation holds.

In Figure~\ref{fig:emittedener} , we estimate $\overline{E}_{\alpha}$ in the 0.5--2 a.u.
velocity range, or equivalently from 15 to 50 eV, for all the targets.
Our results agree with the experimental findings~\cite{surdutovic2018}. 
The dependence of the mean energy value with the projectile charge $Z$ 
is surprisingly sensible, which can duplicate the proton results. 
This effect can be attributed to the depletion caused by the 
multicharged ions to the yields of low energy electrons. In the high 
electron energy regime, the CDW falls on the simple first Born 
approximation, surviving the $Z^{2}$ law. Then, the ratio in 
Eq.~(\ref{40}) cancels out and $\overline{E}_{\alpha}$ becomes a 
universal value independent on $Z$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Emitted electron angles}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

As mentioned before, secondary electrons contribute to the total damage. 
Then, not only is the ejection energy important but also the angle 
of emission. Once again, we can consider the single differential cross 
section in terms of the ejected electron solid angle $\Omega$, 
$d\sigma_{\alpha,nl}/d\Omega$, to be expressed as a distribution function, 
and the mean angle $\overline{\theta}_{\alpha}$ can be defined as
\begin{equation}
\overline{\theta}_{\alpha}=\frac{\langle\theta_{\alpha}\rangle}
{\langle 1\rangle}=\frac{1}{\sigma_{\alpha}}\sum\limits_{nl}
\int d\Omega\,\theta\,\frac{d\sigma_{\alpha,nl}}{d\Omega}
\end{equation}

\begin{figure}[t!]
\centering
\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/Fig6.eps}
\caption{.}
\label{fig:emittedang}
\end{figure}

Figure~\ref{fig:emittedang} shows $\overline{\theta}_{\alpha}$ for our six elements of 
interest. A new important dependence of $\overline{\theta}_{\alpha}$ 
with $Z$ is observed. It is a general belief~\cite{Rudd1992} that the 
angular dispersion of emitted electrons are nearly isotropic. This effect 
is caused by the insignificant angular anisotropy of sub-50-eV yield. 
A typical value for the ejection angle considered in the literature is 
$\overline{\theta}_{\alpha}\sim 70\degree$~\cite{surdutovic2018}, and 
it is quite correct in the range of validity of the first Born 
approximation for any target. But, when a distorted wave approximation 
is used, $\overline{\theta}_{\alpha}$ decreases substantially with $Z$ 
in the intermediate energy region, as observed in Figure~\ref{fig:emittedang}. 
For example, for C$^{+6}$ impact, the Bragg peak occurs at 0.3 MeV/amu, 
where $\overline{\theta}_{\alpha}$, computed with the CDW method, is 
about half of the value obtained with the first Born approximation. 
This correction should close the damage to the forward direction.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A modified stoichiometric model}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\bibitem{rahman2016}
M. A. Rahman and E. Krishnakumar,
Electron ionization of DNA bases,
J. Chem. Phys. \textbf{144}, 161102 (2016).

\bibitem{mozejko2003}
P. Mozejko and L. Sanche, 
Cross section calculations for electron scattering from DNA and RNA bases.
Radiat Environ. Biophys \textbf{42}, 201 (2003).
